---
title: "Breast Cancer 2"
author BCM Team: "Franklin Ajisongun, Matilda Wysocki, Nelson Subervi, Malia Kniwght"
date: "3/9/2019"
output: html_document
---

```{r}
library(ggplot2)
```


load the data 

--str() can be used to view the struction of the object, as well as variable type -- head(data, n) provides the first n rows of the data frume

```{r}
# read csv
breast.cancer.raw <- read.csv("data.csv")

# View first 6 rows
head(breast.cancer.raw)
```

Replace spaces and underscores in variables names with an underscore
```{r}
names(breast.cancer.raw)<- gsub(" ", ".", names(breast.cancer.raw))
names(breast.cancer.raw)<- gsub("_", ".", names(breast.cancer.raw))
```

Order Columns by Alpha
```{r}
breast.cancer.raw<- breast.cancer.raw[,order(names(breast.cancer.raw))]
```

Moving ID and Diagnosis
```{r}
ID<- c("diagnosis", "id")

for (i in 1:2){
  breast.cancer.raw<- breast.cancer.raw[, c(ID[i], setdiff(names(breast.cancer.raw), ID[i]))]
}
```
Remove unnecessary abject
```{r}
rm(ID, i)
```

Mean and Five Number Summary for each variable
```{r}
variable.summary <- data.frame(do.call(cbind, lapply(breast.cancer.raw[c(-1,-2)], summary)))
```


```{r}
variable.summary
```

Quick count of cancer rate in data set
```{r}
table(breast.cancer.raw$diagnosis)
```

Looking at distribution for area.mean variable

```{r}
hist(breast.cancer.raw$area.mean, 
     main = "Distribution of cell Area Means",
     xlab = "Mean Area",
     col = 'green')
```


Add race and age varibles. I created relative probabilities of risk by risk by race. 
I just guessed at age probabilities, increasing as patients gets older: https://ww5.komen.org/BreastCancer/RaceampEthnicity.html

```{r}
# Set seed to your favorite number for replication
set.seed(8675309)

# Divide datasets by diagnosis
breast.cancer.raw.M <- breast.cancer.raw[which(breast.cancer.raw$diagnosis == "M"),]
breast.cancer.raw.B <- breast.cancer.raw[which(breast.cancer.raw$diagnosis == "B"),]

# Assign risk probabilities by race
breast.cancer.raw.M$race <- sample( c("White", "Black", "Asian", "Hispanic", "Other"), 
                                    nrow(breast.cancer.raw.M), 
                                    replace = TRUE,
                                    prob = c(0.41, 0.31, 0.11, 0.14, 0.03))


breast.cancer.raw.B$race<- sample( c("White", "Black", "Asian", "Hispanic", "Other"), 
                                nrow(breast.cancer.raw.B), 
                                replace = TRUE,
                                prob = c(0.28, 0.28, 0.18, 0.20, 0.06))

# Assign risk probabilities by age

breast.cancer.raw.M$age <- sample( 18:40, 
                                   size = nrow(breast.cancer.raw.M), 
                                   replace = TRUE, 
                                   prob = c(0.005, 0.005, 0.006, 0.006, 0.009, 0.012, 0.016, 0.022, 
                                           0.025, 0.08, 0.17, 0.19, 0.14, 0.044, 0.038, 0.029, 0.027,
                                           0.024, 0.021, 0.028, 0.042, 0.03, 0.031) )
breast.cancer.raw.B$age <- sample( 18:40, 
                                   size = nrow(breast.cancer.raw.B), 
                                   replace = TRUE, 
                                   prob = c(0.01, 0.01, 0.01, 0.012, 0.015, 0.018, 0.022, 0.032,
                                            0.04, 0.1, 0.2, 0.201, 0.12, 0.04, 0.033, 0.027, 0.022,
                                            0.018, 0.016, 0.014, 0.02, 0.01, 0.01) )

# Combine tables back together
breast.cancer<- rbind(breast.cancer.raw.M, breast.cancer.raw.B)

# Moving variables

ID<- c("diagnosis", "age", "race", "id")

for( i in (1:4)){
  breast.cancer <- breast.cancer[, c(ID[i], setdiff(names(breast.cancer), ID[i]))]
}

# Delete all unneeded data
rm(breast.cancer.raw.M, breast.cancer.raw.B, breast.cancer.raw, ID, i)
```

display the data
```{r}
head(breast.cancer)
```

```{r}
# Convert variables to factor
breast.cancer$diagnosis<- as.factor(breast.cancer$diagnosis)
breast.cancer$diagnosis<- as.factor(breast.cancer$diagnosis)

# Some methods will require a numeric binary input
breast.cancer$diagnosis.1M.0B<- ifelse(breast.cancer$diagnosis == "M", 1, 0)

# Create function for normalization
normalize<- function(x){
  return((x-min(x))/(max(x)-min(x)))
}

# Normalize variables
breast.cancer.n <- as.data.frame(lapply(breast.cancer[5:34], normalize))

# Name new variables with a suffix '.n'
colnames(breast.cancer.n)<- paste0( colnames(breast.cancer.n), ".n")

# Order columns by Alpha
breast.cancer.n<- breast.cancer.n[, order(names(breast.cancer.n))]

# Pull numreic variables into list for correlation calculations
variables <- breast.cancer[!(names(breast.cancer) %in% c("id","diagnosis","race"))]

# Order Columns by Alpha
breast.cancer.n<- breast.cancer.n[, order(names(breast.cancer.n))]

# Combine new variables with old
#breast.cancer.n<- breast.cancer.n[, order(names(breast.cancer.n))]

# Combine new variables with old
breast.cancer <- cbind(breast.cancer, breast.cancer.n)

# Moving variables
ID <- c("diagnosis.1M.0B", "diagnosis", "age", "race", "id")

for (i in 1:5){
  breast.cancer<- breast.cancer[, c(ID[i], setdiff(names(breast.cancer), ID[i]))]
}

# Remove unnecessary objects
rm(breast.cancer.n, normalize, ID, i)
```

look a little closer at some of these variables

```{r}
# Display row percentages
prop.table(table(breast.cancer$race, breast.cancer$diagnosis), 1)
```

```{r}
# Display column percentages
prop.table(table(breast.cancer$race, breast.cancer$diagnosis), 2)
```

```{r}
# Display % Malignat by Age
age.diagnosis <- as.data.frame(prop.table(table( age = breast.cancer$age,diagnosis = breast.cancer$diagnosis), 1))

head(age.diagnosis)
```

```{r}
age.diagnosis<- age.diagnosis[age.diagnosis$diagnosis == 'M',]
```

```{r}
#windowsFlots( Times = windowsFont("Time New Roman"))
ggplot2::ggplot(age.diagnosis, aes(Freq, x = age, group = 1)) +
   geom_point(aes(y = Freq, x = age, group = 1),
              shape = 1, data = age.diagnosis) +
   geom_smooth(method = lm) +
   theme(legend.position = "bottom", 
         legend.direction = "horizontal",
         legend.title = element_blank()) +
  labs( x = "Age in Years",
        y = "Percentage of Malignance") + 
  ggtitle("Percentage of Malignant Patients of Age") + 
  scale_color_manual(values = fill) + 
  theme(axis.line = element_line(size=1,colour = "black"), 
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), 
        panel.border = element_blank(),
        panel.background = element_blank()) + 
   theme(plot.title = element_text(family = "Times"), 
        text = element_text(family = "Times"),
        axis.text.x = element_text(colour = "black", size = 9),
        axis.text.y = element_text(colour = "black", size = 9),
        legend.key = element_rect(fill = "white", colour = "white"),
        plot.margin = unit(c(1, 1, 1, 1), "in")) 
  
 
```

```{r}

ggplot(breast.cancer, aes(x=breast.cancer$perimeter.worst,
                          group=breast.cancer$diagnosis,
                          fill=breast.cancer$diagnosis)) +
       geom_histogram(position="identity",binwidth=10, alpha = .5) + 
      theme_bw() + 
      xlab("Worst Perimeter") + 
      ylab("Number of Patients") + 
      ggtitle("Distribution of Cell Perimeter by Diagnosis Group") + 
      guides(fill=guide_legend(title="Benign/Malignant"))

```

```{r}
library(corrplot)
# Create Correlaton plot
variables<- variables[,-33]
variables_corr <- variables[1:32]
corrplot(cor(variables_corr), method = "circle", order = "AOE")
```

```{r}
# Remove unnecessary objects
rm(age.diagnosis, variables, variables_corr )
```

split Data into ~ 70% train ~ 30% test 
```{r}
# Assign variable for Train/Test designation
breast.cancer$train_0_test_1 <- sample(0:1, 
                                       nrow(breast.cancer), 
                                       replace = TRUE, 
                                       prob = c(0.7,0.3))

# Create Train/Test datasets
bc.train <- breast.cancer[ breast.cancer$train_0_test_1 == 0,]
bc.test <-  breast.cancer[ breast.cancer$train_0_test_1 == 1,]

# Create result datasets
results.train <- bc.train[c("id", "race", "age", "diagnosis")]
results.test <-  bc.test[c("id", "race", "age", "diagnosis")]
```

Recursive partitioning is a quick and easy way to predict categorial varibles. However, the rpart proceduce in R may be biased toward poly level categorial variables

```{r}
model1 <- diagnosis ~ radius.mean + texture.mean + perimeter.mean +
                      area.mean +smoothness.mean + compactness.mean +
                      concavity.mean +concave.points.mean + symmetry.mean +
                      fractal.dimension.mean + radius.se + texture.se +
                      perimeter.se + area.se + smoothness.se +compactness.se +
                      concavity.se + concave.points.se + symmetry.se+ 
                      fractal.dimension.se + radius.worst + texture.worst +
                      perimeter.worst + area.worst + smoothness.worst +
                      compactness.worst + concavity.worst +concave.points.worst +
                      symmetry.worst + fractal.dimension.worst
```

```{r}
library(rpart)
library(rpart.plot)

# Train the model using the model above, not including race and age
rec.part1<- rpart( formula = model1, method = "class", data = bc.train)

# Plots the rp map
rpart.plot(rec.part1)
```

```{r}
# Displays Model Spec
printcp(rec.part1)
```


```{r}
# Output prediction for training set
pred1<- predict( rec.part1, type = 'class')

# Output predictions for test set
pred.new1 <- predict(rec.part1, newdata = bc.test, type = 'class')

# Confusion Matrices for train set
table(pred1, bc.train$diagnosis)
```

```{r}
# Confusion Matrices for test set
table(pred.new1, bc.test$diagnosis)
```

```{r}
# Basic Model with all cel measure, race and age added
model2 <- diagnosis ~ radius.mean + texture.mean + perimeter.mean + area.mean +
                      smoothness.mean + compactness.mean + concavity.mean +
                      concave.points.mean + symmetry.mean +fractal.dimension.mean+
                      radius.se + texture.se + perimeter.se + area.se + 
                      smoothness.se + compactness.se+concavity.se +
                      concave.points.se + symmetry.se + fractal.dimension.se +
                      radius.worst + texture.worst + perimeter.worst + area.worst +
                      smoothness.worst + compactness.worst + concavity.worst + 
                      concave.points.worst + symmetry.worst +
                      fractal.dimension.worst + race + age
```

```{r}
# Train the model using the model above, including race and age
rec.part2<- rpart( formula = model2, method = 'class', data = bc.train)

# Plots the rp map
rpart.plot(rec.part2)
```
```{r}
# Displays Model Specs
printcp(rec.part2)
```
```{r}
# Output predictions for training set
pred2 = predict(rec.part2, type = 'class')

# Output predictions for test set
pred.new2 = predict(rec.part2, newdata=bc.test, type = 'class')

# Confusion Matrices for Train/Test
table(pred2, bc.train$diagnosis)
table(pred.new2, bc.test$diagnosis)
```

```{r}
str(breast.cancer)
```
