---
title: "BCM Team"
output: 
  flexdashboard::flex_dashboard:
    orientation: row
    vertical_layout: fill
---

```{r setup, include=FALSE}
library(flexdashboard)
library(knitr)
#library(DT)
#library(rpivotTable)
library(ggplot2)
library(plotly)
library(dplyr)
#library(openintro)
library(highcharter)
#library(ggvis)
```


```{r}
data<- read.csv("BreastCancer_DATA.csv")
```

```{r}
mycolors<- c("blue", "#FFC125", "darkgreen", "darkorange")
```

Interactive Data Visualation
=======================================

colume
---------------------------------------

### Breast Cancer Analysis
```{r}
valueBox(paste("Tumor"),
         color = "warning")
```

### Patients in US

```{r}
valueBox(length(data$race),
         icon = "fa-user")
```

### **Malignan**

```{r}
gauge( sum(data$diagnosis == "M"),
             min = 0,
             max = 569,
             gaugeSectors(success = c(0,150),
                          warning = c(150,240),
                          danger = c(240,569),
                          colors = c('green', 'yellow', 'red')))
```

### White

```{r}
valueBox( sum(data$race == "White"),
          icon = 'fa-building' )
```

### African - America
```{r}
valueBox( sum(data$race == "Black"),
          icon = 'fa-building'
          )
```

### Hispanic
```{r}
valueBox( sum(data$race == "Hispanic"),
          icon = "fa-building")
```

### Asian
```{r}
valueBox( sum(data$race == "Asian"),
          icon = "fa-building")
```

Row
----------------------------

### Plot1

```{r}
p2 <-  ggplot( data, aes( x= race, fill = diagnosis)) + 
          geom_bar( stat = "count") +
          labs(title = "Race: Benign/Malignant") +
          guides(fill=guide_legend(title="Benign/Malignant"))
p2.2 <- ggplotly(p2)
p2.2
```


### Plot 2

```{r}
p2 <-  ggplot( data, aes( x= age, fill = diagnosis)) + 
          geom_bar( stat = "count") +
          labs(title = "Age: Benign/Malignant") +
          guides(fill=guide_legend(title="Benign/Malignant"))
p2.2 <- ggplotly(p2)
p2.2
```

Map
=================================================


```{r}
data <- read.csv("USCS_OverviewMap.csv")

p5 <- data %>%
          select(State = Area, Population) %>%
            mutate( State = as.character(State))

highchart() %>%
            hc_title( text = "Breast Cancer in US",
                      style = list(color = "blue",
                                   useHTML = TRUE)) %>%
            hc_subtitle( text = "Soure: BreastCance.csv") %>%
            hc_add_series_map( usgeojson,
                               p5,
                               name = "State", 
                               value = "Population", 
                               joinBy = c("woename", "State")) %>%
           hc_mapNavigation( enabled = TRUE)
```

Data exploration
==============================================



Row
-------------------------

### Distribution of Cell Perimeter 

```{r}
data<- read.csv("BreastCancer_DATA.csv")

breast.cancer <- data

ggplot(breast.cancer, aes(x=breast.cancer$perimeter.worst,
                          group=breast.cancer$diagnosis,
                          fill=breast.cancer$diagnosis)) +
       geom_histogram(position="identity",binwidth=10, alpha = .5) + 
      theme_bw() + 
      xlab("Worst Perimeter") + 
      ylab("Number of Patients") + 
      ggtitle("Distribution of Cell Perimeter by Diagnosis Group") + 
      guides(fill=guide_legend(title="Benign/Malignant"))
```

### Distribution of Cell Area

```{r}
ggplot(breast.cancer, aes(x=breast.cancer$area.worst,
                          group=breast.cancer$diagnosis,
                          fill=breast.cancer$diagnosis)) +
       geom_density(position="identity",binwidth=10, alpha = .5) + 
      theme_bw() + 
      xlab("Area") + 
      ylab("Frequency") + 
      ggtitle("Distribution of Area by Diagnosis Group") + 
      guides(fill=guide_legend(title="Benign/Malignant"))
```



Row
-------------------------

### Distribution of Cell Compactness


```{r}
ggplot(breast.cancer, aes(x=breast.cancer$compactness.worst,
                          group=breast.cancer$diagnosis,
                          fill=breast.cancer$diagnosis)) +
       geom_density(position="identity",binwidth=10, alpha = .5) + 
      theme_bw() + 
      xlab("Area") + 
      ylab("Frequency") + 
      ggtitle("Distribution of Compactness by Diagnosis Group") + 
      guides(fill=guide_legend(title="Benign/Malignant"))
```


### Distribution of Cell Radius Worst 


```{r}
ggplot(breast.cancer, aes(x=breast.cancer$radius.mean,
                          group=breast.cancer$diagnosis,
                          fill=breast.cancer$diagnosis)) +
       geom_density(position="identity",binwidth=10, alpha = .5) + 
      theme_bw() + 
      xlab("Area") + 
      ylab("Frequency") + 
      ggtitle("Distribution of Area by Diagnosis Group") + 
      guides(fill=guide_legend(title="Benign/Malignant"))
```


Kmeans Model
=========================

Column {.sidebar data-width=500}
---------------


### Clustering with features

```{r}
Breast_Cancer_data <- read.csv("data.csv")
Breast_Cancer_data$X<- NULL
Breast_Cancer_data<- Breast_Cancer_data[,-1]
Breast_Cancer_data$diagnosis<- factor(ifelse( Breast_Cancer_data$diagnosis == "B", "Benign", "Malignant"))

library(factoextra)
Breast_Cancer_data2<- transform(Breast_Cancer_data)

all_pca <- prcomp(Breast_Cancer_data2[,-1], cor = TRUE, scale = TRUE)

fviz_pca_biplot(all_pca, 
                col.ind = Breast_Cancer_data$diagnosis, 
                col = "black", 
                palette = "jco",
                geom = "point",
                repel = TRUE,
                legend.title = "Diagnosis",
                addEllipses = TRUE)
```

```{r}
nrows <- NROW(Breast_Cancer_data)

set.seed(218)

index <- sample(1: nrows, 0.7 * nrows)

train <- Breast_Cancer_data[index,]
test <- Breast_Cancer_data[-index,]
```



### Cluster plot

```{r}
predict.kmeans<- function(newdata, object){
  centers <- object$centers
  n_centers <- nrow(centers)
  dist_mat <- as.matrix(dist(rbind(centers, newdata)))
  dist_mat<- dist_mat[-seq(n_centers), seq(n_centers)]
  max.col(-dist_mat)
}

learn_Kmeans <- kmeans(train[,-1], centers = 2)
library(factoextra)
learn_Kmeans$cluster<- ifelse(learn_Kmeans$cluster == 2, "Benign", "Malignant")
fviz_cluster( learn_Kmeans, data = train[,-1])
```


Column 
----------
###Confusion Matrix 

```{r}
library(e1071)
library(caret)

predict.kmeans<- function(newdata, object){
  centers <- object$centers
  n_centers <- nrow(centers)
  dist_mat <- as.matrix(dist(rbind(centers, newdata)))
  dist_mat<- dist_mat[-seq(n_centers), seq(n_centers)]
  max.col(-dist_mat)
}

learn_Kmeans <- kmeans(train[,-1], centers = 2)

pre_kmeans <- predict.kmeans(test[,-1], learn_Kmeans)
pre_kmeans<- as.factor(ifelse(pre_kmeans == 2 ,"Benign", "Malignant"))
cm_kmeans <- confusionMatrix(pre_kmeans, test$diagnosis)
cm_kmeans
```

SVM Model 
=========================

### Confusion Matrix


```{r}
library(e1071)
learn_svm <- svm( diagnosis~., data = train)
pre_svm <- predict(learn_svm, test[, -1])
cm_svm<- confusionMatrix(pre_svm, test$diagnosis)
cm_svm
```


Prediction: SVM model
===================

```{r}
patient<- read.csv("data.csv", header = TRUE, stringsAsFactors = FALSE)
patient$X<- NULL
```

**John: Patient number 20th**
```{r}
John<- patient[20,]   ## 20th patient
John[,c(1,2)]         ## Benign
```

**Mary: Patient number 19th**
```{r}
Mary <- patient[19,]    ## 19th patient
Mary[,c(1,2)]   
```


```{r}

gamma <- seq(0, 0.1, 0.005)
cost <- 2^(0:5)
parms <- expand.grid(cost= cost, gamma = gamma) # 231

acc_test<- numeric()
accuracy1<- NULL
accuracy2<- NULL
  
for(i in 1:NROW(parms)){
  learn_svm<- svm( diagnosis~., data = train, gamma = parms$gamma[i], cost = parms$cost[i])
  pre_svm <- predict(learn_svm, test[,-1])
  accuracy1<- confusionMatrix(pre_svm, test$diagnosis)
  accuracy2[i]<- accuracy1$overall[1]
}

acc <- data.frame( p = seq(1,NROW(parms)), cnt = accuracy2)


opt_p <- subset(acc, cnt == max(cnt))[1,]
sub <- paste("Optimal number of premeter is", opt_p$p, "(accuracy:", opt_p$cnt,") in C5.0")

learn_imp_svm <- svm(diagnosis~., data = train, cost = parms$cost[opt_p$p], gamma = parms$gamma[opt_p$p])
pre_imp_svm<- predict(learn_imp_svm, test[,-1])
cm_imp_svm <- confusionMatrix(pre_imp_svm, test$diagnosis)

# for print output 
cancer_diagnosis_predict_p <- function(new, method = learn_svm){
  new_pre <- predict(method, new[,-1])
  new_res<- as.character(new_pre)
  return(paste("Patient ID:", new[,1], " => Result: ", new_res, sep = ""))
}


# for submission output
cancer_diagnosis_predict_s <- function(new, method = learn_svm){
  new_pre <- predict(method, new[,-1])
  new_res<- as.character(new_pre)
  return(new_res)
}
```

### Model Testing Function (Use only 1 test data)


```{r}
cancer_diagnosis_predict_p(John)
```

```{r}
cancer_diagnosis_predict_p(Mary)
```

### Make Submission Output (Use test dataset)

```{r}
#library(kableExtra)
t <- patient[-index,]
origin<- t$diagnosis
t$diagnosis<- NULL
r <- cancer_diagnosis_predict_s(t)

sub <- data.frame(id = t$id, predict_diagnosis = ifelse( r == "Malignant", "M", "B"), origin_diagnosis = origin)
sub$correct <- ifelse(sub$predict_diagnosis == sub$origin_diagnosis, "True", "False")
kable(head(sub,20))
```